Towards a Universal Facebook of the Ancient World

### 2016-01-24 source/18/EAGLE2016_Full_Paper_Broux-XeLaTEX.tex
Paper 18 (Broux) fixed with the Coptic and Egyptian fonts (using XeLaTeX); file EAGLE2016_Full_Paper_Broux-XeLaTEX.tex is now the master file. The compiled version (Broux-XeLaTEX.pdf) updated to the pdfs folder as submission_18.pdf

%%%%%%%%
%TITLE:%
%%%%%%%%
\title{Towards a Universal Facebook of the Ancient World}
\author[KU]{Yanne Broux\corref{first}}
\address[KU]{KU Leuven / Research Foundation Flanders (FWO)}
\cortext[first]{Corresponding author. Email: Email: yanne.broux@kuleuven.be}
\date{2015-09-16}
\begin{document}

\maketitle
\begin{abstract}
Facebooking the past. The idea grew while developing a database of all people mentioned in texts from Greco-Roman Egypt
(350 BC–AD 800). Thanks to Trismegistos' role in EAGLE, Named Entity Recognition can now be applied to almost 500,000
Latin inscriptions from the Roman Empire, and some 400,000 clusters containing personal names can be extracted. This
collection of names will lead to a large-scale study of naming practices in the ancient world, and how these reflect
changes in society at large.
\end{abstract}
\keywords{Latin epigraphy, Trismegistos, Named Entity Recognition, Social Network Analysis, Onomastics, Prosopography}

@@

Apart from texts, since 2008 Trismegistos also intensively deals with places and people. Building on open access to the
full text in repositories, Named Entity Recognition procedures are used to create lists of toponyms and anthroponyms
that occur in the ancient sources. This was first applied during the socio-onomastic project `Creating Identities in
Graeco-Roman Egypt' (KU Leuven OT project 2008-2012), on a corpus of about 50,000 papyri and ostraca in the Duke
Databank of Documentary Papyri (papyri.info). With the additional support of a Hercules Grant, (`An Interdisciplinary
Database of Proper Names in Late Pharaonic, Graeco-Roman and Byzantine Egypt (ca. 800 BC – AD 640)'; 2008-2014) the
work on the core data could be finished in just over two years, resulting in a database with almost half a million
references to people (Trismegistos People) and an additional hundred-thousand or so place names (Trismegistos Places)
mentioned in texts from Egypt.

On the basis of Trismegistos People, several studies on naming practices, modes of identification and identity issues in
Greco-Roman Egypt have been published. Two PhDs focused on the longstanding tradition of double names: the first dealt
with the Ptolemaic period, when this form of polyonymy served to cross ethnic borders \citep{Coussement}, the
other with the Roman period, when the practice was adapted by the local elite to distinguish themselves from the hoi
polloi and to resemble Roman nomenclature \citep{Broux2015a}. The data of the fourth century AD provided new insights into
the spread of Christianity in Egypt \citep{Depauw2013}. The development of a standardized identification
cluster, consisting of a person's name, patronymic and metronymic \citep{BrouxDepauw}, as well as the use of fixed
expressions to denote illegitimacy \citep{Broux2015b}, name change \citep{Broux2013a} and official identification \citep{Broux2010}, were all related to the legal reorganization of the population under the Romans and the
ensuing tax reform \citep{Broux2013b}. Most recent studies focus on the influence of Roman socio-linguistic practices on
Greek and Egyptian conventions \citep{Depauw} and how network analysis can provide us with new insights regarding
onomastic habits and what they say about cultural identity and social status \citep{Broux2015c}.

@@

Trismegistos opted for a combination of a gazetteer and a rule-based approach. For the gazetteer a three-tiered
onomastic database structure was developed in Trismegistos People, dealing with names, name variants, and declined name
variants respectively. The first database, NAM, currently has 34,106 entries, e.g. the name Isidoros. Each name is
connected to a set of variants in all possible languages. As a rule, only very minor dialectal or orthographical
variation is allowed in the `native' language, in this case Greek (e.g. \begin{otherlanguage*}{greek}
E>is'idwroc
\end{otherlanguage*} and
\begin{otherlanguage*}{greek}
>Is'idoroc
\end{otherlanguage*}). Many of the variants are actually created by renderings of a name in other languages, e.g.
Ỉsytrs, ȝysydwrs or ⲓⲥⲓⲇⲱⲣⲉ in Egyptian and Coptic. In all there
are 177,290 variants in the NAMVAR database. Finally, for each of the variants, the declined forms were created, e.g.
\textgreek{>Is'idwrou (}genitive) or \textgreek{>Is'idwrwi (}dative). This NAMVARCASE database is the largest with
667,677 entries, and this set is used as a gazetteer for NER. It was developed on the basis of the set of names of some
40,000 individuals with titles listed in the Prosopographia Ptolemaica (KU Leuven) and was supplemented with new names
collected during the `Creating Identities' project \citep{Depauw2009}. Through cooperation with the Lexicon of
Greek Personal Names (Oxford University) the names of some 300,000 individuals can now also be integrated.

>>>>

\usepackage{fontspec}

\usepackage{polyglossia}

\setmainlanguage{english}
\setotherlanguage{greek}

\newfontfamily\greekfont[Script=Greek]{CMU Serif}

%\setmainfont{lmr}
%%%%%%%%
%TITLE:%
%%%%%%%%
\title{Towards a Universal Facebook of the Ancient World}
\author[KU]{Yanne Broux\corref{first}}
\address[KU]{KU Leuven / Research Foundation Flanders (FWO)}
\cortext[first]{Corresponding author. Email: Email: yanne.broux@kuleuven.be}
\date{2015-09-16}
\begin{document}

\maketitle

\begin{abstract}
Facebooking the past. The idea grew while developing a database of all people mentioned in texts from Greco-Roman Egypt
(350 BC–AD 800). Thanks to Trismegistos' role in EAGLE, Named Entity Recognition can now be applied to almost 500,000
Latin inscriptions from the Roman Empire, and some 400,000 clusters containing personal names can be extracted. This
collection of names will lead to a large-scale study of naming practices in the ancient world, and how these reflect
changes in society at large.
\end{abstract}
\keywords{Latin epigraphy, Trismegistos, Named Entity Recognition, Social Network Analysis, Onomastics, Prosopography}

@@

Apart from texts, since 2008 Trismegistos also intensively deals with places and people. Building on open access to the
full text in repositories, Named Entity Recognition procedures are used to create lists of toponyms and anthroponyms
that occur in the ancient sources. This was first applied during the socio-onomastic project `Creating Identities in
Graeco-Roman Egypt' (KU Leuven OT project 2008-2012), on a corpus of about 50,000 papyri and ostraca in the Duke
Databank of Documentary Papyri (\url{papyri.info}). With the additional support of a Hercules Grant, (`An Interdisciplinary
Database of Proper Names in Late Pharaonic, Graeco-Roman and Byzantine Egypt (ca. 800 BC – AD 640)'; 2008-2014) the
work on the core data could be finished in just over two years, resulting in a database with almost half a million
references to people (Trismegistos People) and an additional hundred-thousand or so place names (Trismegistos Places)
mentioned in texts from Egypt.

On the basis of Trismegistos People, several studies on naming practices, modes of identification and identity issues in
Greco-Roman Egypt have been published. Two PhDs focused on the longstanding tradition of double names: the first dealt
with the Ptolemaic period, when this form of polyonymy served to cross ethnic borders \citep{Coussement}, the
other with the Roman period, when the practice was adapted by the local elite to distinguish themselves from the \emph{hoi
polloi} and to resemble Roman nomenclature \citep{Broux2015a}. The data of the fourth century AD provided new insights into
the spread of Christianity in Egypt \citep{Depauw2013}. The development of a standardized identification
cluster, consisting of a person's name, patronymic and metronymic \citep{BrouxDepauw}, as well as the use of fixed
expressions to denote illegitimacy \citep{Broux2015b}, name change \citep{Broux2013a} and official identification \citep{Broux2010}, were all related to the legal reorganization of the population under the Romans and the
ensuing tax reform \citep{Broux2013b}. Most recent studies focus on the influence of Roman socio-linguistic practices on
Greek and Egyptian conventions \citep{Depauw} and how network analysis can provide us with new insights regarding
onomastic habits and what they say about cultural identity and social status \citep{Broux2015c}.

@@

Trismegistos opted for a combination of a gazetteer and a rule-based approach. For the gazetteer a three-tiered
onomastic database structure was developed in Trismegistos People, dealing with names, name variants, and declined name
variants respectively. The first database, NAM, currently has 34,106 entries, e.g. the name Isidoros. Each name is
connected to a set of variants in all possible languages. As a rule, only very minor dialectal or orthographical
variation is allowed in the `native' language, in this case Greek (e.g. \textgreek{Εἰσίδωρος} and \textgreek{Ἰσίδορος}). Many of the variants are actually created by renderings of a name in other languages, e.g. %{\fontspec{Liberation Sans}ȝysydwrs}
Ỉsytrs, {\fontspec{Gentium}ȝysydwrs} or {\fontspec{ArialCoptic} ⲓⲥⲓⲇⲱⲣⲉ} in Egyptian and Coptic. In all there
are 177,290 variants in the NAMVAR database. Finally, for each of the variants, the declined forms were created, e.g.
\textgreek{Ἰσίδωρου}) or \textgreek{Ἰσίδωρωι} (dative). This NAMVARCASE database is the largest with
667,677 entries, and this set is used as a gazetteer for NER. It was developed on the basis of the set of names of some
40,000 individuals with titles listed in the Prosopographia Ptolemaica (KU Leuven) and was supplemented with new names
collected during the `Creating Identities' project \citep{Depauw2009}. Through cooperation with the Lexicon of
Greek Personal Names (Oxford University) the names of some 300,000 individuals can now also be integrated.
